%%% ACKNOWLEDGEMENTS
%%% optional argument in \begin{acknowledge}[...] gives style of title

%\chapter*{Acknowledgements}
\begin{acknowledge}[\normalfont\large\bfseries]

This thesis has been the culmination of over ten years of study into the area of data synchronization for mobile devices, started during my undergraduate years at Brock University in St. Catharines, Ontario, through my time working for IBM, and with the OpenTAPAS Project, until today -- sometimes with support, other times purely of my own volition.  As such, there are a number of people who rightfully must be acknowledged here, for without their aid, this body of work may have never come to pass.

Firstly, I'd like to acknowledge Prof. Jon Radue of Brock University in St. Catharines, Ontario, who acted as my undergraduate supervisor in what became \emph{The jSyncManager}.  Without the jSyncManager, I wouldn't have been writing this thesis today, as its genesis is firmly rooted in the success of this project.

Secondly, I'd like to thank Dr. Morgan Price, who initiated what became the OpenTAPAS Project.  It was through Morgan's need for a data synchronization solution that led him to me, and to eventually bring me on board the OpenTAPAS Project as their mobile device data synchronization expert.  We had great ideas and a great team, all of which was brought down by politics outside our control.

It was through Dr. Price that I was introduced to Dr. Jens Weber, who would become my graduate supervisor, and an indispensable hand in completing this thesis.  Dr. Weber has been the lens through which the ideas presented in this thesis were able to come into better focus.  It was Dr. Weber who stood up for me when I decided to go back to school to complete my Masters degree, and for this I will forever be grateful.

No acknowledgement is complete without a nod to the family support structure behind the researcher.  As such, I'd like to thank my mother and father, Pauline and Robert Richard (Dick) Barclay for their love and support throughout my life.  This work is every bit as much a testament to their efforts as it is to my own.  Special acknowledgements go out to my brother and only sibling, David Barclay, not only for the life-long bond between us that is the equal of no other, but for helping me design the outline for the Mobile Salesperson A user model.  Dave, you may never have the interest or opportunity to explore education in the manner that I have, but you've made your mark on science, and it will forever be recorded in these pages.

Finally, extra special acknowledgements go out to my wife, G\"ulcan.  My love, a phase in our lives is about to pass, and I look forward to nothing more than exploring all of our remaining ones together.  You have been my rock throughout the writing of this thesis, which I herein dedicate whole-heartedly to you.

\end{acknowledge}
