\chapter{Letter of Implied Consent}\label{impliedconsent}

\begin{center}
\textbf{Optimizing Synchronization Cost for Mobile Devices:The Expedient Trickle Sync Algorithm}
\end{center}

You are invited to participate in a study entitled Real-World Mobile Device Synchronization Frequency in Commercial Applications that is being conducted by Brad BARCLAY.Brad Barclay is a Graduate Student in the Department of Computer Science at the University of Victoria and you may contact him if you have further questions via e-mail at yaztromo@uvic.ca.As a graduate student, I am  required to conduct research as part of the requirements for a degree in Computer Science. It is being conducted under the supervision of Dr. Jens H. WEBER. You may contact my supervisor at jens@uvic.ca. 

\begin{flushleft}\textbf{Purpose and Objectives}
\end{flushleft}The purpose of this research project is to determine the norms and best practices used by organizations when synchronizing handheld computing devices (mobile phones, Palm-style handheld computers, etc.) to corporate databases.   \begin{flushleft}\textbf{Importance of this Research}\end{flushleft}
Research of this type is important because such information is currently lacking, and can be of great use to researchers who are attempting to find better and more cost-effective mechanisms for conducting the synchronization of handheld computing devices. \begin{flushleft}\textbf{Participants Selection}\end{flushleft}
You are being asked to participate in this study because you are/were a known user of the jSyncManager, a Free Open Source software package for synchronizing PalmOS-based handheld devices. \begin{flushleft}\textbf{What is involved}\end{flushleft}
If you agree to voluntarily participate in this research, your participation will include answering a simple 5-minute web-based survey.  The questions in the survey may, at your option, require a brief analysis of your existing databases and/or log files used or generated when synchronizing mobile devices within your organization.

\begin{flushleft}\textbf{Inconvenience}
\end{flushleft}Participation in this study may cause some inconvenience to you, including the brief amount of time required to respond to this survey. \begin{flushleft}\textbf{Risks}\end{flushleft}
There are no known or anticipated risks to you by participating in this research.  \begin{flushleft}\textbf{Benefits}\end{flushleft}
The potential benefits of your participation in this research include more accurate models on how existing organizations handle the process of synchronizing information from portable, semi-connected devices, which will allow the researcher to evaluate more cost-effective protocols and best practices for performing such synchronization tasks in the future. \begin{flushleft}\textbf{Voluntary Participation}\end{flushleft}
Your participation in this research must be completely voluntary. If you do decide to participate, you may withdraw at any time without any consequences or any explanation. If you do withdraw from the study your data will be impossible to remove once submitted, as no records of what data is attached to which respondent will be retained. \begin{flushleft}\textbf{Anonymity}\end{flushleft}
In terms of protecting your anonymity no identifying information will be requested or permitted.  No digitally identifying information (such as MAC addresses or IP addresses) will be recorded, stored, or retained as part of the survey submission process.  \begin{flushleft}\textbf{Confidentiality}\end{flushleft}
Your confidentiality and the confidentiality of the data will be protected by the lack of any identifying information being requested, stored, or retained. All information will be stored within an encrypted database on a machine stored in a locked office within the University of VictoriaŐs Computer Science Department. \begin{flushleft}\textbf{Dissemination of Results}\end{flushleft}
It is anticipated that the results of this study will be shared with others in the following ways:  via published articles, thesis, and via presentations.  The final study results will be made available to participants upon request.\begin{flushleft}\textbf{Disposal of Data}\end{flushleft}
Data from this study will not be disposed of.  The anonymous survey responses will be recorded in an appendix of the researcherŐs thesis.  The electronic database used to collect and store the initial results will be deleted using a 7-pass erase function at the completion of this study. \begin{flushleft}\textbf{Contacts}\end{flushleft}
Individuals that may be contacted regarding this study include Mr. Brad BARCLAY $<$yaztromo@cs.uvic.ca$>$ and Dr. Jens H. WEBER $<$jens@uvic.ca$>$. In addition, you may verify the ethical approval of this study, or raise any concerns you might have, by contacting the Human Research Ethics Office at the University of Victoria (250-472-4545 or ethics@uvic.ca).  By completing and submitting the questionnaire, YOUR FREE AND INFORMED CONSENT IS IMPLIED and indicates that you understand the above conditions of participation in this study and that you have had the opportunity to have your questions answered by the researchers.  \begin{center}\textbf{\textit{Please retain a copy of this letter for your reference.}}\end{center}