\chapter{Survey Invitation to Participate}\label{surveyinvitation}
(Sent to invitees via e-mail).

You are invited to participate in a study entitled Real-World Mobile Device Synchronization Frequency in Commercial Applications that is being conducted by Brad BARCLAY.Brad Barclay is a Graduate Student in the Department of Computer Science at the University of Victoria.  As a graduate student, I am  required to conduct research as part of the requirements for a degree in Computer Science.The purpose of this study is to determine the norms and best practices used by organizations when synchronizing handheld computing devices (mobile phones, Palm-style handheld computers, etc.) to corporate databases.  This research is important; the lack of existing public research in this area prevents the comparison of different practices, both real and hypothesized, in an attempt to improve overall cost-effectiveness of data synchronization for mobile devices.The survey consists of seven questions, and will take no more than five minutes to complete.  No identifying information of any type will be requested, stored, or retained, including the names of participants, organizations they belong to, IP addresses, or digital signatures.To participate, you are invited to visit the the survey website at [WEB ADDRESS OF SURVEY SITE].  You are invited to take a moment to read the Letter of Information for Implied Consent, and at your option to complete the short survey.Please note that participation in this study is VOLUNTARY.  Invitees should not feel obligated to participate based on professional affiliation.If you have any questions pertaining to this survey, you are invited to contact the researcher, Brad BARCLAY, via e-mail at $<$yaztromo@cs.uvic.ca$>$.Your time and consideration are greatly appreciated.
\begin{flushleft}Brad BARCLAY\\M.Sc. Candidate,\\Department of Computer Science,\\University of Victoria,\\Victoria, BC, CANADA\end{flushleft}
