\chapter{User Model Study Questionnaire}\label{modelquestions}


\begin{enumerate}\item During regular business hours, how many times on average do you expect users to synchronize their mobile devices against your organizations database(s)?\item Considering your answer to question 1, which of the following best describes how you arrived at your response? (check one):
\begin{itemize}\item Organization mandated best practice\item Result from computerized synchronization logs or database query\item Approximate guess based on experience with users
\end{itemize}\item During non-business hours (evenings, weekends, holidays, etc.) how many times per day do you expect users to synchronize their mobile devices against your organizations database(s)?\item Considering your answer to question 2, which of the following best describes how you arrived at your response? (check one):
\begin{itemize}\item Organization mandated best practice\item Result from computerized synchronization logs or database query\item Approximate guess based on experience with users
\end{itemize}\item Approximately how many records are contained within the database table(s) your users synchronize their mobile devices against?\item During regular business hours, how many times on average are the records within this(/these) database table(s) expected to be modified (either through the synchronization of modified records, or through other means) per day?\item Considering your answer to question 6, which of the following best describes how you arrived at your response? (check one):
\begin{itemize}\item Result from computerized synchronization logs or database query\item Approximate guess based on experience with the database(s) in question.
\end{itemize}
\end{enumerate}
