%%%
\chapter{Simulation Environment}\label{simulation_chapter}
%%%

\section{Introduction}

The purpose of the simulation environment is to provide a consistent mechanism for evaluating the cost and efficiency of various synchronization protocols.  It takes a series of configuration parameters as input which describe the characteristics of the environment we wish to simulate along with how long to run the simulation, and outputs a linear equation expressing the average cost per day for the synchronization algorithm in question.

In this chapter, we'll explore the environmental elements being simulated, the mathematical models used to drive the simulation, how the simulation logic is controlled, the protocol used to communicate with synchronization adaptors, along with a discussion of the functionality of the three synchronization protocols developed to compare the ETS protocols effectiveness against.

The simulation itself runs at a resolution of one-second intervals.  We define a \emph{tick} to be a simulated one second interval within the simulation.

\section{Simulated Elements}\label{sim_elements}

The goal of the simulation environment is to evaluate the overall cost of a synchronization algorithm by following a model user through a series of days as they\footnote{For the purposes of this paper, \emph{they} is used as a gender-neutral identifier for a single individual, as supported by references \cite{theyref3, theyref2, theyref1}.} synchronize their wireless mobile device through a programmed series of locations.  Within each location there may be a variety of network types available, each with their own cost structure and data transfer rates.  These elements will each be presented in this section.

\subsection{Networks}\label{networks}

As we are evaluating the synchronization costs related to wireless mobile devices, we require some form of network transport for the transmission of synchronization data and metadata within our simulation environment.  These are expressed as a series of network objects, one per network which our sample user may intersect during the simulated day.

The networks each have three primary attributes inside the simulation:  their transfer rate expressed as bytes per second ($N_{rate}$), the cost associated with each byte transferred ($N_{cost}$), and the probability per second that the connection can be lost ($P_{connection lost}(N)$).  Each of these elements is an important part of the simulation execution, as they permit the simulation to determine the duration of a series of network operations, their related monetary cost, and at which probability a connection may be lost during a synchronization session.

\subsection{Locations}

A location is a specific geo-physical region that the user may be within during their simulated day.  A location is simply defined by zero or more networks which are accessible within that location.  Note that a given network may exist within multiple locations; for example, it is expected that generally a cellular network will be available in most locations within most urban areas.  Thus, the relationship between locations and networks is an N:M relationship.

\subsection{Locales}\label{locales}

Locales build upon the physical locations by layering atop them temporal information; thus a locale is a location along with a time in which that location is entered.  This is used to permit the simulation to change from one location to the next during the duration of the simulated day.

Along with the time in which a location is entered ($L_{entry}$), the locale also provides information on both the frequency of synchronization ($L_{sync frequency}$), and the number of expected mobile database accesses ($L_{expected accesses}$).  These are described internally as the expected number of synchronizations per second, and the expected number of accesses per second.  Note that we typically do not expect a user to make multiple synchronization requests nor to have multiple database accesses within a a single second; as such these values are expected to be represented by small positive fractions.

Through the use of locale objects, we can permit the user to enter the same physical location at different times of day, with different mobile device usage patterns.  For example, a user who telecommutes from home and who uses their mobile device frequently during the normal work day may have a high expected database accesses value during business hours, but would have an exceedingly low expected database accesses at night while asleep.  In such a case the location is static, but the users usage patterns change.  Thus, the relationship between locations and locales is 1:N.  As the days are simulated from time 00:00:00 until time 23:59:59, the first defined locale is automatically entered at 00:00:00, and the last defined local persists until 23:59:59.

\subsection{Records}\label{records}

A record is the smallest unit of data exchanged during the synchronization process, and represents a logical unit of compound information.  For the purposes of the simulation, we do not actually model the record or its data, but instead simply maintain a record identification number, size, and revision number.  The identification number ($R_{ID}$) is used as a matching mechanism between a record stored in the mobile database and a record stored within the server database.  The size ($R_{size}$) is used for calculating how long it will take to transfer the record, as well as how much its transfer will cost.  Finally, the revision number ($R_{revision}$) is used as a mechanism for comparing matching records within the mobile and server databases to determine differences between the records.

The revision number is of particular importance, as it is used in two different contexts within the simulation environment.  Firstly, it is used during a mobile database access by the simulated user to assess whether to increment the intangible cost factor which relates to the use of out-of-date information; if the revision number of the mobile database record $M_k$ is less than that of the matching record on the server database $S_k$, the intangible cost factor is incremented by the difference $S_k - M_k$.  Secondly, it is used to provide the synchronization protocol implementation with a list of records requiring synchronization, returning the set of all records with identification number $k$ where $S_k > M_k$.

While it is possible for the record size to be specified manually, the record size is typically assigned during the simulations set-up stage randomly based on a Normal distribution (\ref{gaussian}).  In this way the records can be generated with varying sizes, with a pre-set mean and variance.

Records also have an implicit probability of access by the user of the mobile device.  The simulation assumes that records are not evenly used, but that instead certain records are accessed more frequently than others.  This is intended to model typical real-world experience with databases.  While this probability is not stored within the record itself, it is based on a records ID number and the total number of records within the database containing it.  This will be discussed further in sub-section \ref{databases}.

\subsection{Databases}\label{databases}

A database is a logical collection of records.  For the sake of the simulation environment, databases are always of a fixed size.  The system maintains two separate simulated databases, one for the server, and one for the mobile device.  Ideally, for the intangible cost of synchronization to be zero, these two databases should be identical at all times.  However, in the typical situation, it is expected that these two databases will in fact differ for the bulk of the simulation run \cite{1704823}.

The server database $S$ maintains an expected number of modification arrivals per second and a maximum number of simultaneous modification arrivals.  These modifications reflect changes made to the database by other database clients, exclusive of our simulated mobile device, but inclusive of all other mobile devices within the organization that synchronize against the same database.  Using a Poisson Process (\ref{poisson}), a series of probabilities are calculated for each possible arrival value from $0$ to $S_{max arrivals}$ inclusive, using the expected number of modifications per second.  During every simulated second, a random value is selected and compared against this series of probabilities to determine how many records should be modified during the current simulation time.  Once a number $v$ between $0$ and $S_{max arrivals}$ has been selected based on the calculated probabilities, $v$ records within the database are selected for modification.  These $v$ records are selected at random using a Normal distribution (\ref{gaussian}), such that the record with ID $S_{max}/2$ has the highest probability of selection, and the records with ID $0$ and $S_{max}$ have a very low probability of selection, with all other values having a probability falling upon a Normal curve.  Once each record to be modified has been identified, its revision number $R_{revision}$ is incremented by one.

The mobile database $M$ handles whether or not the user has viewed a record during each one second interval in the simulation, based upon the expected mobile database accesses value stored in the current locale object.  During every tick, we choose a random number $n$ and compare it to the expected mobile database access rate per second in $L_{expected accesses}$.  If $L_{expected accesses} >= n$, an access has occurred.  We then determine which record has been accessed during this interval in a manner identical to that of how we determine which record to modify in the server database $S$, using a Normal distribution.  In this way the most-modfied records on the server are also the most frequently accessed records on the mobile database.  Unlike the server database, however, mobile database accesses are limited to one per tick, and as such form a Bernoulli Process (\ref{bernoulli}).

\subsection{User}\label{user}

Tying the above simulated elements together is the user object ($U$).  Only a single simulated user exists within a simulation run, and they persist throughout every simulated day.  The user object maintains the list of locales available to the user, and manages the transition between one locale and another during the simulated day.  As such, the user object maintains a 1:N relationship with the locale objects.  The user object also contains the single instance of the mobile database, and as such has a 1:1 relationship with the mobile database object.

\section{Models}

Manipulation of the simulated elements is governed by a series of mathematical models, which dictate the behaviour of the simulation process by ensuring the simulated elements remain within realistic bounds.  The models transform pseudo-random numbers into values within a defined range in order to simulate unknown information that falls within a known (or expected) set of bounds \cite{RefWorks:38, RefWorks:40}.  This section details these models as they are utilized within the simulation environment, along with the rationale for their selection.

\subsection{Pseudo-Random Number Generation}\label{random_nums}

A key ingredient to simulating discrete, random events is the reliable generation of random numeric values \cite{RefWorks:38}.  For simulation, it is generally desirable to generate values which are as close to uniformly distributed within the interval $(0, 1)$ as possible.  As it is impossible to generate random values using purely deterministic techniques \cite{RefWorks:38}, we must choose a pseudo-random number generator (PRNG) algorithm which produces uniformly distributed numeric values, with an extremely long period between repetitions.  Additionally, the pseudo-random number generation algorithm should produce random values very quickly, as an algorithm which requires a large number of instructions may cause our simulation to require significant runtime to arrive at a result, as a single simulation run may require tens of millions random numbers (indeed, based on experimentation a one year simulation can easily require in excess of 30 million random values).  Thus, a balance between the appearance of randomness, long period, equidistribution, and computational efficiency is the ideal trait of a pseudo-random number generator for use in our simulator \cite{wiki:mersennetwister}.

The PRNG selected for the simulation is the Mersenne Twister algorithm \cite{RefWorks:43}, as it meets all of these requirements.  Its period before values start to repeat themselves has been proven to be $2^{19937}-1$, has a very high order of equidistribution, and passes numerous statistical tests for randomness.  It is also quite fast.  In testing during PRNG evaluation for the simulator, the LGPL Mersenne Twist implementation used \cite{mersennetwistcode} was slightly more efficient than the Mac OS X built-in random number generators \texttt{rand} and \texttt{drand48}.  The fact that the efficiency is not worse than the platforms random number generator, coupled with its improved equidistribution and significantly longer period make it an excellent choice for discrete event simulation \cite{RefWorks:43}.

\subsection{Normal Distribution}\label{gaussian}

Characterizing the types of data that can be synchronized in a generalized manner is nearly impossible;  the types of information that can be synchronized are as varied as any other form of computer data.  In addition, without studying a specific and widely adopted synchronized database, it is nearly impossible to generalize the access frequency of the records within the database.

Due to these factors, a normal distribution is used for calculating both the size of records, and to determine which records are accessed and/or modified by the client and server.  In the case of the record sizes, a mean size in kilobytes is provided, and record size values are calculated based upon a normal distribution around this mean.  By default, to create records within a reasonable size for a handheld device, a mean of 10 KB is used, with a variance of 5.  This ensures that within two standard deviations of the mean, approximately $95\%$ of records will be within the range $[5..15]$ KB.

During the simulation run, both the server and the user are in frequent need to select a record for viewing or modification.  As the individual records are assumed to be accessed at different frequencies, with some records being accessed more frequently than others, a normal distribution is also applied to them.  In this case, the mean is one-half the database size.  A record is selected by calculating a random normally distributed value, converting it into the percent possibility of its selection in the cumulative density function, and then multiplying this percentage by the total number of records.  In this manner, records around the mean will be selected more often, while records around the ends of the database are selected less frequently.

The rationale behind using the normal distribution in both cases revolves around the central limit theorem \cite{RefWorks:38, RefWorks:40}.  According to the central limit theorem, if the sum of the identically distributed random variates have a finite variance, it will be approximately normally distributed \cite{RefWorks:38, wiki:CLT}.  As we have no fixed distribution to affix our simulation to, and as we intend to simulate a generalized set of records, the normal distribution is a suitable and reasonable substitute.

Generation of random, standard normally distributed values (with a mean of $0.0$ and a standard deviation of $1.0$ ($\varphi_{0, 1}$)) is accomplished via the polar form of the Box-Muller transformation, as presented in \cite{RefWorks:42}.  This transformation generates a pair of random, normally distributed values from a single uniformly distributed random value (generated by the Mersenne Twister algorithm \cite{RefWorks:43, mersennetwistcode}).  During each iteration of the random normal value method, a pair of values is calculated, with one cached.  During a subsequent call to the method, the cached value is returned.

Generation of random, normally distributed values with other means ($\mu$) and standard deviations ($\sigma$) is also supported.  Calculation of these values $\varphi_{\mu,\sigma^2}$ is based upon a random, standard normally distributed value $\varphi_{0, 1}$, which is then transformed using the following equation \cite{RefWorks:38}:
\begin{equation}
\varphi_{\mu,\sigma^2} = \mu + \sigma \varphi_{0,1}\label{normal_dist}
\end{equation}

\subsection{Bernoulli Process}\label{bernoulli}

For testing whether or not a discrete event occurs at a given time $t$, a Bernoulli Process is used \cite{RefWorks:40}.  During simulation set-up, an expected number of events per hour ($E_{hour}$) is read from the configuration file, and divided to determine a probability-per-second for the event ($P(n)$), such that:
\begin{equation}
P_{bernoulli}(1)=\frac{E_{hour}}{3600}\label{bernoulli_prob}
\end{equation}

Every simulated second, an event trial occurs by selecting a uniformly distributed random number from the PRNG, and is compared against $P_{bernoulli}(1)$.  If the value is less than or equal to $P_{bernoulli}(1)$, the event occurs.  If it is less than $P_{bernoulli}(1)$, the event does not occur.  Note that due to this event determination mechanism, values of ($E_{hour}>3600$) are not supported.  However, this limitation is not an issue for the events where the Bernouilli Process is used, as sub-second accuracy is not provided, and values in excess of $3600$ would require an average of more than one event per second.  In cases where multiple events per second is desirable, a Poisson Process (\ref{poisson}) is used instead.

Bernoulli Process events are used for events which either occur or not during a one second interval.  Typically this is used to simulate user access to a handheld database record, as it is assumed the user may only access a single record at once.  Every locale has an independent ($E_{hour}$) for handheld record accesses.

\subsection{Poisson Process}\label{poisson}

For events which may occur multiple times in a single tick, a Poisson Process is utilized \cite{RefWorks:38, RefWorks:40, RefWorks:32}.  During simulation set-up, an expected number of events per hour ($E_{hour}$) is provided, along with a maximum number of permitted events per interval ($E_{max}$).  From these two values, a series of cumulative event probabilities is calculated, in the range of $[0..E_{max}]$.  That is, the probability of exactly $n$ events occurring during an interval with arrival rate $r$ and for interval $t$ is calculated as:
\begin{equation}
P_{poisson}(n) = \sum_{k=0}^n \frac{(rt)^k}{k!}e^{-rt}\label{poisson1}
\end{equation}

Every simulated second, an Poisson trial occurs by selecting a uniformly distributed random number $Z$ from the PRNG, and finding the smallest $n$ such that:
\begin{equation}
P(n)>Z\label{poissontrial}
\end{equation}
The result is the number of events which occurred during the current second.

Poisson Process events are used to simulate those elements which can have a variable number of occurrences per interval.  In particular, this type of event is used when simulating modifications to the server database (as we assume this database has multiple clients, and thus may see multiple modifications simultaneously).

\section{Controllers}

Tying together the simulated elements and the mathematical models for the simulation are a set of simulation controllers which manage the execution of the simulation, modification of the data elements, application of the models, and the recording of the results.

\subsection{Time Controller}\label{time_controller}

An important aspect of the simulation is the flow of time.  Time has a significant impact on the entire simulation execution; all of the event-based models are evaluated at fixed time intervals.  In addition, locale switching is time-based.  Thus, the proper simulation and use of these elements requires some form of time controller.

The time controller as implemented in the simulation environment is relatively simple.  It uses a 1 second resolution counter, which runs each day from $0$ to $86400$ within a loop.  Within the loop, two types of possible events are evaluated:  \emph{tick} events, and \emph{alarm} events.  Both are implemented using a registration and call-back system, along with a protocol that registered routines must implement to satisfy the call-back subsystem.

Tick events are used for simulated elements that require evaluation within regular, one-second intervals.  These include all simulated elements using either a Bernoulli Process (\ref{bernoulli}) or a Poisson Process (\ref{poisson}), along with any elements that use their own distribution or process which must be evaluated at regular intervals that cannot be pre-computed.  During every iteration of the tick loop within the time controller, every registered tick listener has its \texttt{activateTick} method called, passing with it the current tick value.  Note that tick listeners are not guaranteed to be called in any specific order.

Alarm events are used for simulated elements which desire an event to occur at a specific, specified tick during the simulated day.  There are used by elements which know in advance when they would like to be activated, such as locale switching.  During every tick iteration, any alarms registered waiting for the current tick value have their \texttt{activateAlarm} methods called, in no particular order.  In order to enhance the simulation run, all alarm listeners are stored within a a series of arrays, each of which is associated with a specific alarm time.  These arrays are themselves referenced from within a hash hashing the tick time against the array of listeners for that time.  Only those tick values which have an associated alarm as represented by an array, minimizing the memory requirements and minimizing the search time required to test if any alarms need to be activated for any given tick.

The time controller itself is only capable of executing a single simulated day.  For multi-day simulations, the time controller's \texttt{run} method is executed in sequence once per desired simulated day by the simulation controller, as discussed in subsection \ref{simcontroller}.

\subsection{Cost Controller}\label{cost_controller}

The entire purpose of the simulation environment is to generate a result; the value of this result is managed by and stored within the cost controller.  The cost controller for the simulation keeps track of the cost of accessing out-of-date information on the handheld device, the cost of using network resources, as well as the total data transferred within the simulation session.  These three values are stored as running totals during the duration of the simulation run, with the final results presented as the mean average result over the total number of simulated days.

Modification of the real cost value within the cost controller is typically managed by the Synchronization Controller (\ref{synccontroller}), which handles the simulation of the synchronization session and calculates the cost to synchronize a given record through a specified network.  Modification of the intangible cost of using out-of-date information is handled by the mobile database, which intercepts read requests against the database, and calculates the distance of the handheld record's revision from that of the matching server record's revision, adding the result to the running total for the intangible cost within the cost controller.

The cost controller can output its results as a linear equation of the form:
\begin{equation}
y=mx+b\label{lineareq}
\end{equation}
where $y$ is the average cost, $m$ is the average intangible disappointment factor, $x$ is the unknown intangible disappointment unit cost, and $b$ is the average real network access cost.  Note that as this is a standard linear equation, $m$ is also the slope of the linear equation.  When analyzing the results, a lower slope $m$ indicates less financial risk associated with using out-of-date data, and a lower constant $b$ indicates a lower average network cost.  Better synchronization algorithms must thus strive to decrease both elements.

\subsection{Synchronization Controller}\label{synccontroller}

The synchronization controller handles tasks on behalf of the different synchronization protocols.  It encapsulates those functions which are of common need to most of the synchronization adapters (\ref{protocols}), in particular, recording the initiation of a synchronization session (\ref{connection_and_auth}), reporting the set of differences between the server database and the handheld database (\ref{id_mod_recs}), synchronizing a single record (\ref{writing_records}), and handling the disconnection (\ref{disconnection}).  Other requirements for a proper synchronization solution are not implemented, in significant part because there is no actual record data within the simulation.

There are two important features of note within the record synchronization routine.  Firstly, the single record synchronization routine permits for the correct recording of cost information associated with a given network.  This ensures consistency when comparing the results from different protocols, as every protocol must synchronize each record through the synchronization controller.  Secondly, it permits us to simulate random disconnections from each network, based on the disconnection probability attached to each.  This is calculated by evaluating how many seconds it is expected to take to transfer each record, calculating the probability of a disconnection within this time period, retrieving a random variate, and comparing it to the calculated probability.  If the variate is less than the probability, a disconnection occurs, the synchronization protocol is immediate informed, and no other record synchronizations are permitted until a new synchronization session is established.

The synchronization controller calculates the duration required to synchronize the records in a single synchronization session.  As all synchronization occurs within a singe second in simulation time, this value is used to ensure that another synchronization session cannot be started when the previous session is logically in progress.

\subsection{Simulation Controller}\label{simcontroller}

The simulation controller ties the entire simulation engine together by providing the facility to load the simulation parameters from a file, and the permitting a multi-day simulation to occur.  It uses an instance of the time controller (\ref{time_controller}) to run multiple single-day simulations, up to a user-specified maximum number of days.  Additionally, the simulation controller permits the automated execution of a batch of these multi-day simulations with each individual session within the batch having its own uniquely generated server and handheld databases.

As part of its duties, the synchronization controller resets the time controller for each new day, and calculates and maintains the values required to calculate the variance ($\sigma$) of the cost from day-to-day.  The variance is calculated using an algorithm derived from formulas described by Knuth in \cite{RefWorks:42}, section 4.2.2, as formulas 15 and 16.  This algorithm has been enhanced to work with linear equations instead of real-numbered values, so as to generate results appropriate to the simulators solutions.

The simulation controller supports a call-back mechanism, for alerting an implementing application when individual days complete.  This is used in the simulation front-end to update a progress bar, alerting the user that work is being accomplished.

\section{Comparative Protocols}\label{protocols}

In order to successfully evaluate the Expedient Trickle Sync algorithm, a series of algorithms have been implemented for the sake of comparison.  Two algorithms are implemented as an experimental control:  the null synchronizer which produces a worst-case cost scenario by never synchronizing any data, and a random data synchronizer.  In addition, two algorithms based on existing industry-standard synchronization mechanisms, Fast Sync \cite{RefWorks:2, RefWorks:22, RefWorks:31} and Slow Sync \cite{RefWorks:1, RefWorks:2, RefWorks:24, RefWorks:31}.  All four algorithms are presented within this section.

\subsection{Null Synchronizer}

The Null Synchronizer algorithm is an algorithm which provides no synchronization capabilities, providing a worst-case scenario for the intangible cost, while simultaneously providing a best-case scenario for the real cost.  Lacking any synchronization features, records quickly become out-of-date, with the situation getting worse as time progresses.  However, with the lack of synchronization is also the lack of use of any network resources to accomplish the synchronization.  As such, the Null Synchronizer exists to provide a worst-case control value; no synchronization solution should ever have a result with a slope as steep as that for the null synchronizer.

Lacking any synchronization facilities, the null synchronizer models the real-world situation where a given user never synchronizes their mobile device with an associated centralized server, but where the centralized server sees frequent modification to the same information used by the mobile device user.

\subsection{Random Synchronizer}

The ransom synchronization algorithm doesn't model any real-world usage scenarios, and is simply provided as a control value.  It synchronizes at random based on an expected number of synchronizations per hour within each distinct locale (as do the Slow Synchronizer (\ref{slowsync}) and Fast Synchronizer (\ref{fastsync}) algorithms), however during synchronization the random synchronizer selects a uniformly distributed random number of records to synchronize ($n$) from within the range $[0..M_{size}]$, and then selects $n$ records from the mobile database at random to synchronize.  It completely ignores whether or not the records selected even require synchronization, and thus may partially synchronize data which does not require it.  In areas where multiple networks are available, one is selected at random for use during the synchronization session.

As the random variate $n$ used is uniformly distributed, over a large number of synchronizations the expected average number of records synchronized per session should be $M_{size}/2$.

\subsection{Slow Synchronizer}\label{slowsync}

The slow synchronizer simulates a set of real-world synchronization protocols, such as the Slow Sync mode of Palm Computing's HotSync Protocol \cite{RefWorks:1, RefWorks:2, RefWorks:24, RefWorks:31}.  It is designed to read every record from the handheld database during every synchronization, determine on the server-side if any differences occur, and if so update the record and send it back to the mobile device.  The slow sync algorithm is discussed further in \ref{id_mod_recs}.

In the simulator implementation, the slow synchronizer algorithm transfers every record in the mobile database across the network, with every modified record transferred twice.  This mimics the real network transmission needs of slow sync style protocols, which transmit every record to the server, which then compares them against its database, and transmits back any records which have been modified.

The slow synchronizer algorithm synchronizes at random based on the expected number of synchronizations per hour within each distinct locale.  This mimics the fact that slow sync based handhelds typically require the user to initiate the synchronization session.

\subsection{Fast Synchronizer}\label{fastsync}

The fast synchronizer algorithm simulates another set of real-world algorithms, such as the Fast Sync mode of Palm Computing's HotSync Protocol \cite{RefWorks:2, RefWorks:22, RefWorks:31}, Apple's iSync \cite{RefWorks:33}, and others.  It is the most frequently used synchronization algorithm.  Like the slow synchronization algorithm, it synchronizes at random based on the expected number of synchronizations per hour within each distinct locale.  However, unlike the slow synchronizer algorithm, the fast synchronization algorithm is able to determine which records require synchronization, and only transmits the records which do across the network.  As such, the rate of synchronization should be identical to the slow synchronizer algorithm, however the actual data transferred should be significantly reduced, due to the fact that only a subset of the records are transmitted during each synchronization session.  The fast synchronization algorithm is discussed further in \ref{id_mod_recs}.

\section{Limitations}

While the simulation environment is designed to closely detail the realities of mobile data synchronization, it is imperfect, facing a number of limitations.  However, it is felt that these limitations do not have a significant impact on the simulation results.  These limitations will be discusses herein.

Firstly, all synchronization performed is unidirectional; in effect we're always updating the handheld device based on changes generated on the server, but never vice-versa.  However, as a bidirectional synchronization can be composed of two unidirectional synchronizations in opposing directions, and as this limitation affects all the simulated synchronization algorithms equally, it is felt that the impact on the results is negligible.  Indeed, assuming the server database's expected modifications per hour rate includes expected modifications from the simulated mobile device, the impact on the results will be virtually nil.

Another issue in the simulation environment is that the cost associated with manual conflict resolution is complete ignored.  As the simulation environment doesn't simulate user-initiated changes to the mobile database, and as it contains no real data, no conflicts occur.  It can be argued that for databases where conflict resolution can be completed in a completely automated manner, however, that the cost of conflict management will be virtually nil, and thus can be easily ignored.

The mobile users daily behaviour within the simulation is assumed to be completely static.  While this is not realistic, it has been known for some time that mobile users exhibit various degrees of regularity in their behaviour patterns \cite{1704823}, and thus this is not felt to be a significant issue.

Finally, no attention is paid in the cost calculation to mechanisms to reduce the amount of information transmitted in an update packet; the simulation assumes that the entire modified record is transmitted to the out-of-date database for synchronization.  While most existing synchronization algorithms do nothing to reduce this data size, the use of data compression and/or delta packets (as described in \ref{update_packet}) can be used to equally reduce the size of packets for any of the synchronization algorithms used (with the notable exception of the null synchronizer, as it lacks any synchronization facilities).  Thus, for the sake of comparison, data reduction methods can be safely ignored within the simulation.

None of these limitations are felt to have a significant impact on the final results; the end results continue to permit differing algorithms to be accurately compared to each other.  Removing these limitations would significantly impact the performance of the simulation, without impacting upon the results in a significant or meaningful way.  Thus it is felt that these limitations are sufficiently insignificant, and thus have little real impact on the usefulness of the final results.
