%%%
\chapter{User Model Study} \label{proposed_study}
%%%

\section{Purpose}

Accurate evaluation of the various synchronization algorithms within the simulation environment is dependant upon a set of accurate user models.  These models dictate the frequency of record modification, data synchronization, and the users daily movements throughout their environment.  An accurate model will thus provide a more useful result for each algorithm, and thus a more accurate cost comparison can be made.

Building accurate user models requires information on the habits and patterns of users with handheld mobile devices which are synchronized against a database.  Unfortunately, such habits and patterns are lacking within the existing literature.  In addition, due to the fast evolution experienced in the mobile device arena, any possible existing studies older than one or two years may not reflect the habits of today's mobile device user.  As such, a study of mobile device user habits was required in order to build a series of accurate, real-world user models to evaluate the simulated synchronization algorithms.

\section{Methodology}

In order to study the habits of the target population for the purpose of creating accurate user models, we created a series of seven survey questions.  These questions are listed in Appendix \ref{modelquestions}.  The questions are designed to determine the average number of times users are expected to synchronize during both work hours and off-work hours, along with the size of the database being synchronized against and the average number of changes to said database during an average day.  Four of the questions are designed to elicit specific numerical responses from which to design a user model, while the remaining three questions are designed to determine the respondent's confidence level for the values reported in three of the other four questions.

The survey itself is presented as a web application, using LimeSurvey \cite{limesurvey}.  This online survey tool was installed on a secure server at the University of Victoria, and was used to manage the survey invitations, provide the facility for the target population to complete the survey via the World-Wide Web, and to anonymize and store the survey results.  Prior to completing the survey, the users are invited to download and read a \underline{Letter of Implied Consent}, which is available in Appendix \ref{impliedconsent}.

Participants were self-selected, with invitations being sent to organizations known or believed to use data synchronization software.  This was achieved by targeting Internet mailing lists whose purpose is the discussion of implementing mobile data synchronization solutions in multi-user organizations.  Organizations known to use the jSyncManager \cite{RefWorks:30} were also invited to participate.  A copy of the invitation letter can be found in Appendix \ref{surveyinvitation}.  In order to minimize responses from organizations or individuals outside the target population, each invitation contained a survey-software generated token, which must be supplied in order to respond to the survey.  Each token may only be used once, and no information is stored or retained to attach a specific invitee to a specific token.

The study was reviewed and approved by the University of Victoria Human Research Ethics Board.  A copy of the Certificate of Approval can be found in Appendix \ref{ethicscert}.

\section{Results}

The survey was an abject failure, failing to garner a single response.  The following reasons have been identified for the failure of this project, based on what few responses were received from the initial Letter of Invitation (see Appendix \ref{surveyinvitation}):
\begin{description}
\item[Requested Information Confidential and/or Proprietary:] At least one organization approached felt that the information requested is either confidential and/or proprietary in nature.  While not indicating whether they had suitable information to respond to this survey or not, respondents in this category felt that this information on their synchronization habits may give them a competitive advantage, and thus were unwilling to share it, even when they were informed that only average/aggregate values were going to be requested, and that the survey would be completely anonymous.
\item[Unable to Find Person with Required Knowledge:] In virtually every case, finding the people responsible for handling what is typically a back-end process for most organizations is virtually impossible; publicly advertised contacts within organizations believed to use synchronization systems with mobile devices often didn't know whom to direct the survey to.  Many requests appear to have been lost within organizational bureaucracy, never to be resolved.  Invitations sent to public Internet mailing lists directed towards implementers of existing data synchronization solutions went unanswered; it is possible that those who are in charge of implementing such solutions do not have sufficient authority or interest in participating in such a survey.
\item[No Response:] The largest group of invitees fell into this category, having given no response.  No response was received from any of the targeted mailing list invitations to participate.  The lack of response could be related to any of the above reasons, due to simple lack of interest, due to the feeling that participants wouldn't sufficiently benefit from the research, due to a lack of incentive, or because the selected target mailing lists lack sufficient members within the target audience.
\end{description}

\section{Conclusion}

While the lack of response to the survey was disappointing, it is fortunately not fatal to the research.  Without suitable real-world information on data synchronization patterns, we must rely on invented models which appear plausible, in order to evaluate the synchronization algorithms previously described in \ref{protocols}.  These invented models are presented in \ref{user_profile_data}, along with arguments for their design.

The presence of real-world synchronization patterns would eliminate the possibility of bias in the selection of user model patterns for evaluating the ETS algorithm, and would continue to be of benefit to future research in this area. As such I would like to recommend that future research into the area of data synchronization for mobile devices should consider undertaking this sort of survey again, perhaps improving upon it through the participation of someone with more insider contact and knowledge within the mobile data synchronization implementation community, or through the use of some form of respondent incentive.
