%%%
\chapter{Conclusions}\label{Conc}
%%%

As evidenced by the results, the ETS algorithm shows promise in reducing the cost of data synchronization over the existing Fast Sync and Slow Sync protocols, however it does demonstrate some problem areas which we identify as possible areas of future research.  To conclude this thesis, ideas for future research are discussed, along with my final conclusions.

\section{Future Research}\label{future_research}

There are a variety of areas in which the ETS algorithm can be improved, to provide additional and more consistent cost savings across a variety of different user profile types.  There are also a number of existing limitations with the simulation environment, and as such the application of the ETS algorithm's results in this thesis cannot be extrapolated to real-world users.

Firstly, the current simulation environment continually runs the same day pattern repeatedly, with fixed locale entry times that never differ, for the entirety of the simulation run.  This rigid pattern does not map particularly well to real-world users, where entry and exit times may be variable, and where they exhibit different patterns on different days.  While the ETS algorithm was designed to take some of these issues into account, they were not tested in this thesis due to this limitation in the simulation environment.  Further research and development is needed to improve the simulation environment, to make it capable of handling something closer to real-world user patterns.

Similarly, the current simulation also presumes that day-to-day patterns for the last seven days are generally the same \ref{when_to_sync}.  However, certain classes of users may exhibit different patterns on different days of the week, but which are comparable for the same day of the week.  That is, a given user may not have the same daily pattern, but may have a reasonably comparable pattern on Wednesdays.  Experiments using the ETS algorithm, extended to record several weeks worth of daily access, with the access probabilities based on those of the same day of the week is another avenue of further research.

As mentioned in \ref{limitations}, the results of the experiments are only demonstrably valid for the simple mathematical models selected.  Further experiments using other distributions and arrival models, particularly those with more complexity, would be worthwhile.  In addition to experiments using other models and distributions within the simulation environment, expanding and completing the user model study is key to improving the evaluation of not only the ETS algorithm, but of other potential intelligent synchronization algorithms.  Additionally, it may be possible to use existing wireless data sets to derive accurate user models, such as the data sets available from Dartmouth University's CRAWDAD project \cite{crawdad, 511359, 1160109}.

In terms of improvements to the ETS algorithm itself, there are a variety of areas available for further research.  Firstly, the random outliers visible in several of the 100-day result graphs (figures \ref{simple_a_results_graph-100}, \ref{simple_b_results_graph-100}, \ref{business_a_results_graph-100}, \ref{business_b_results_graph-100}, \ref{sales_a_results_graph-100}, and \ref{trucker_a_results_graph-100}) is of concern; in particular, in the current implementation of the ETS algorithm these random outliers could cause the algorithm to make a poor cost-improvement choice.  As the ETS algorithm appears to converge quickly based on the results of the 100-day experiments, three potential solutions to this issue are readily apparent.  In the first, it should be possible to smooth out the effects of these outliers somewhat by running the same threshold for multiple days, and basing the threshold update decision on the average of these results.  While this would presumably bring any set of results for a single threshold value closer to the norm whenever an outlier occurs, the outlier could still have a significant impact on cost.  A second solution would be to again run the same threshold for multiple days, selecting the median value as the cost for evaluation.  A third solution would be to blend aspects of the first two solutions together, such that when evaluating the results of a multi-day threshold test, an equal number of high and low outliers are removed from the result set, with the remaining results averaged together to come up with a single cost value, which can then be evaluated.

Another area worthy of further examination is the initial threshold point selection.  During development, it was noted at one point that the threshold initialization algorithm (\ref{threshold_initialization}) was only being run during the first simulation run; subsequent runs were not engaging this algorithm.  For some of the tests (such as Simple User A), the results were significantly better than the results presented in this thesis, whereas for other tests (such as Simple User B) the results were significantly worse (indeed, the bug was detected when it was noted that the cost of synchronization for this model was nearly as bad as that for the Null Sync algorithm, as no synchronization was occurring).  While the results with the threshold initialization algorithm were still good for Simple User A, they didn't approach the low cost of the results without this algorithm.  As such, further research can be done on the threshold initialization algorithm.

Another research avenue related to initial threshold point selection would be to extend the threshold calculation to include the calculated thresholds from other devices in the synchronization cloud. In essence, the calculate threshold values could themselves be synchronized through a centralized database, mapping between users with similar habits.  This could decrease the time the ETS algorithm spends attempting to optimize the cost, resulting in a lower average daily cost when using the ETS algorithm.

The arbitrary selection of only one synchronization per 15 minute interval may be unsuitable for many user models.  This value was selected so as to minimize the synchronizations impact on device power, and to prevent a scenario where a low threshold, necessary to synchronize when in a higher-cost network zone, could cause constant synchronization when in a virtually free network zone.  It was deemed desirable when creating the ETS algorithm to avoid this situation, however it could be that a lower interval size would permit the ETS algorithm to decrease costs through more frequent synchronization.  Further research into this area is encouraged.

Next, it is possible that a single threshold value is insufficient for reducing the cost of synchronization.  It may be that a better algorithm could be constructed using the principals of the ETS algorithm, but where each day is broken into zones with different threshold values.  These zones may have fixed start times and durations, or they may have variable start times and durations based on historical observations of the users patterns.

Finally, while section \ref{threshold_problems} contains numerous graphs comparing possible threshold values to expected cost (\ref{threshold_problem_a1}, \ref{threshold_problem_a}, \ref{threshold_problem_b}, \ref{threshold_problem_c}, and \ref{threshold_problem_d}), it should be noted that no such graph has been created for any of the user profiles.  An understanding of how all the different threshold values compare in a specified user profile would permit a better analysis of design needs for such a protocol, and would be of benefit to this research.  A fixed simulated day could be generated and re-run with a variety of possible threshold values to generate this graph, however the reason why it was not undertaken for this thesis is that the computational complexity involved in adequately computing the values for such a graph.  As described in equation \ref{threshold_init}, the initial threshold value is set to $8 \cdot 10^{14}$.  Even accounting only for positive integer threshold values, computing this many values is a significant computational undertaking, requiring a distributed grid of computers to calculate the possibilities within reasonable time constraints.  As a guide, on an Intel Core 2 Duo running at 2.16Ghz, the 250 year simulations took roughly 5 hours to complete, or approximately 18 250 simulated days per hour.  At this rate, $8 \cdot 10^{14}$ simulations would take over 43 million hours to compute.  Distribution over a sufficiently large array of systems, along with only evaluating the threshold values at fixed intervals would permit the creation of such a graph in reasonable time, and is suggested as a very useful avenue of further research.

\section{Concluding Remarks}

Mobile devices continue to proliferate in our society.  During the writing of this thesis, Apple Inc. released their second-generation mobile telephony device, the iPhone 3G.  Their competition likewise has not stood still, and both businesses and consumers have been lining up to purchase these newer devices, which emphasize higher speed data communication features and more local storage alongside the traditional voice and local directory features.  The possibilities for the types of mobile applications that can be used are likewise expanding with better hardware and more ubiquitous and higher speed network connectivity.  Unfortunately, at the same time data prices over cellular networks continue to be excessively expensive, particularly here in Canada, causing many businesses and consumers to try to find ways of minimizing their monthly data transfer to reduce their cellular service bills.  This reduces the utility of these devices, and acts as an economic damper against the kinds of mobile applications which could be developed.  It is for this reality that the ETS algorithm has been designed; by emphasizing overall cost over simply accuracy, a balance can be achieved between both ideals.

As noted in section \ref{future_research}, there is further work to be done in this area, which until now has been under researched.  It is the authors hope that this work will be the precursor to further study, refinement, and innovation in this realm of endeavour.
